---
title: "Random Forest"
output:
   html_notebook:
     code_fold: hide
  
---

```{r setup,include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load fucntions
library(dplyr)
library(ggtern)
library(optmatch)
library(pROC)
source("./Functions.R")

# Global variance for write or not
do.write <- F
# Outputdir
outdir <- "../../Result/RandomForest/NPCpairmatch/"
if(!dir.exists(outdir)){dir.create(outdir)}
if(!dir.exists(paste0(outdir,"model/"))){dir.create(paste0(outdir,"model/"))}
```

# Loading
```{r}
pub.MGS2.prf <- read.table("../../Result/01.Profile/MGS.V2.Caucasian.219.profile")
pub.KO.prf <- read.table("../../Result/01.Profile/Pub_Caucasian/Caucasian.ko.abun.profile")
pub.phe <- read.csv("../../Result/Phenotype/pub.cc.phe.csv")

ko.prf <- read.table("../../Result/01.Profile/KO.profile")
sp.prf <- read.table("../../Result/01.Profile/MGS_V2.profile")
mgs.V2.anno <- read.table("../../Result/01.Profile/00.IGC.MGS.annotation",sep="\t",header=T)
mgs.V2.anno$mgs.V2 <- sub(":",".",mgs.V2.anno$X)
mgs.V2.anno2 <- mgs.V2.anno[,c(8,2,3)]

#FDNA.NS,FDNA.NPC.pick2,FDNA.NPC.pick3
load("../../Result/TimeSeries/NPC/FDNA.NPC.picks.RData")
```

# Cross-validation

Training set came from randomly picked PRs (n=8) and phenotype pair-matched SD and PD.

8 folds for each cross-validation process.

And each try repeat 100 times.

> Cross-validation performed on HPC. Here we just load the results and find the best one.

#decide training set and testing set
```{r}
#Load HLA.mdat
phe.HLA.csv <- read.csv("../../Result/HLA/57_PD1_NPC.2020-4-14.csv")
#phe.HLA.csv$TMB.grp <- ifelse(phe.HLA.csv$TMB>median(phe.HLA.csv$TMB),"high","low")
phe.HLA.csv$TMB.grp <- ifelse(phe.HLA.csv$TMBnew>median(phe.HLA.csv$TMBnew[!is.na(phe.HLA.csv$TMBnew)]),"high","low")
HLA.mdat<-phe.HLA.csv
HLA.mdat.tmblow_HLAE_MT.PID <- as.character((HLA.mdat%>%filter(TMB.grp=="low"&HLAe.SNP=="M/T"))$PID)
#HLA.mdat.tmblow_HLAE_MT.PID <- as.character((HLA.mdat%>%filter(TMB.grp=="low"&HLA_E=="M/T"))$PID)
#HLA.mdat.tmblow_HLAE_MT.PID <- as.character((HLA.mdat%>%filter(TMB.grp=="high"&(HLAe.SNP=="M/M"|HLAe.SNP=="T/T")))$PID)

##
M0.phe <- FDNA.NPC.pick3%>%filter(Ms=="M0"&ATB02==F&!PID%in%HLA.mdat.tmblow_HLAE_MT.PID)
#M0.phe <- FDNA.NPC.pick3%>%filter(Ms=="M0"&ATB02==F&PID%in%HLA.mdat.tmblow_HLAE_MT.PID)
M0.PID <- as.character(M0.phe$PID)
M0.phe$SD2 <- as.factor(ifelse(M0.phe$BE3=="SD","SD","PDPR"))

set.seed(1)
#PR.train.PID <- as.character(sort((M0.phe%>%filter(BE3=="PR"))$PID)[sample(1:12,8)])
PR.train.PID <- as.character(sort((M0.phe%>%filter(BE3=="PR"))$PID)[sample(1:6,4)])

merge.SD.dat <- rbind(
  cbind(set=1,M0.phe%>%filter(PID%in%PR.train.PID)),
  cbind(set=0,M0.phe%>%filter(BE3=="SD")))
#mat.SD <- pairmatch(set~Treatment+Gender+Age+BMI,controls = 2, data=merge.SD.dat)
mat.SD <- pairmatch(set~Treatment+Gender+Age,controls = 2, data=merge.SD.dat)
merge.PD.dat <- rbind(
  cbind(set=1,M0.phe%>%filter(PID%in%PR.train.PID)),
  cbind(set=0,M0.phe%>%filter(BE3=="PD")))
#mat.PD <- pairmatch(set~Treatment+Gender+Age+BMI,controls = 1, data=merge.PD.dat)
mat.PD <- pairmatch(set~Treatment+Gender+Age,controls = 1, data=merge.PD.dat)

merge.dat <- unique(rbind(
  merge.SD.dat[which(!is.na(mat.SD)),],
  merge.PD.dat[which(!is.na(mat.PD)),]))

train.PID <- as.character(merge.dat$PID)


train.phe <- M0.phe%>%filter(PID%in%train.PID)
test.phe  <- M0.phe%>%filter(!PID%in%train.PID)

#
PDPR.train.phe <- M0.phe%>%filter(PID%in%train.PID&BE3!="SD")
PDPR.train.phe$BE3 <- droplevels(PDPR.train.phe$BE3)
PDPR.test.phe  <- M0.phe%>%filter(!PID%in%train.PID&BE3!="SD")

#
method <- 'all'
method <- "cv"

table(train.phe$BE3,train.phe$Treatment)
table(train.phe$BE3,train.phe$Gender)
#ggplot(train.phe,aes(x=Age,y=BMI,color=BE3,shape=Gender)) + 
ggplot(train.phe,aes(x=Age,y=E,color=BE3,shape=Gender)) + 
  geom_point() + stat_ellipse(aes(group=BE3))
```

**function**
```{r}
#
comb.rf.model.pred <- function(rf1,rf2,df=NULL,phe=NULL){
  #makesure the rownames are corresponding to each matrix
  if(is.null(df)){
    pred1 <- predict(rf1,type="prob")
    pred2 <- predict(rf2,type="prob")
    predc <- merge(pred1,pred2,by="row.names",all.x=T)
    for(i in 1:nrow(predc)){
      if(length(which(is.na(predc[i,4:5])))==2){
        predc[i,4:5] <- as.numeric(predc[i,2])/2
      }
    }
    rownames(predc) <- predc$Row.names
    predc <- predc[,-1]
    return(predc)
  }else{
    pred1 <- predict(rf1,df,type="prob")
    pred2 <- predict(rf2,df,type="prob")
    preds <- cbind(SD=pred1[,"SD"],
               raw.PD=pred2[,"PD"],PD=pred1[,"PDPR"]*pred2[,"PD"],
               raw.PR=pred2[,"PR"],PR=pred1[,"PDPR"]*pred2[,"PR"]
               )
    rownames(preds) <- rownames(df)
    if(!is.null(phe)){
      pred.df <- merge(phe,preds,by="row.names")
      auc.PD <- auc(pred.df$BE3=="PD",pred.df$PD,plot=F)
      auc.SD <- auc(pred.df$BE3=="SD",pred.df$SD,plot=F)
      auc.PR <- auc(pred.df$BE3=="PR",pred.df$PR,plot=F)
      auc.df <- rbind(
        data.frame(pred="PD",auc=as.numeric(auc.PD),attr(auc.PD,"roc")[2],attr(auc.PD,"roc")[3]),
        data.frame(pred="SD",auc=as.numeric(auc.SD),attr(auc.SD,"roc")[2],attr(auc.SD,"roc")[3]),
        data.frame(pred="PR",auc=as.numeric(auc.PR),attr(auc.PR,"roc")[2],attr(auc.PR,"roc")[3])
      )
      auc.df <- auc.df[rev(order(auc.df$specificities)),]
      conf.PR <- 1 - mean(rf2$err.rate[,"PR"])
      conf.PD <- 1 - mean(rf2$err.rate[,"PD"])
      conf.SD <- 1 - mean(rf1$err.rate[,"SD"])
      
      return(list(pred=pred.df,auc=auc.df,conf.PR=conf.PR,conf.PD=conf.PD,conf.SD=conf.SD))
    }else{
      return(preds)
    }
  }
}
```

### MGS only

```{r}
#opt1
mgs.rf.tag <- "Ms0.n8_00.MGS.BE3"
#sum.sign.anno <- read.csv("../../Result/TEST/timeSeries/mgs.anova.sum.ms.sgin.anno.csv",row.names = 1)
sum.sign.anno <- read.csv("../../Result/TEST/timeSeries/NPC.mgs.anova.sum.ms.sgin.anno.csv",row.names = 1)
sign.MGS <- as.character(sum.sign.anno$mgs.V2)
#opt2
#mgs.rf.tag <- "MsMd.f8_00.MGS"
mgs.rf.tag <- "MsMd.f8_00.MGS.anova"
mgs.rf.tag <- "MsMd.f8_01.MGS.anova" #45


#phMd.r.sum.sign.anno <- read.csv("../../Result/TEST/timeSeries/mgs.anova.sum.ms.enriched.anno.csv",row.names=1)

phMd.r.sum.sign.anno <- read.csv("../../Result/TEST/timeSeries/NPC.mgs.anova.sum.ms.sgin.anno.csv",row.names=1)
sign.MGS <- as.character(phMd.r.sum.sign.anno$mgs.V2) 

#####
choose.mgs.prf <- sp.prf[sign.MGS,sort(colnames(sp.prf))]
####
mgs.rf.dat <- merge(train.phe[,c("DNAID","SD2")],t(choose.mgs.prf),by.x="DNAID",by.y="row.names")
rownames(mgs.rf.dat) <- mgs.rf.dat$DNAID
mgs.rf.dat <- mgs.rf.dat[,-1]
#DPPR
PDPR.mgs.rf.dat <- merge(data.frame(DNAID=colnames(choose.mgs.prf),t(choose.mgs.prf)),
    PDPR.train.phe[,c("DNAID","BE2")],by="DNAID")
rownames(PDPR.mgs.rf.dat) <- PDPR.mgs.rf.dat$DNAID
PDPR.mgs.rf.dat <- PDPR.mgs.rf.dat[,-1]

#####CV
MGS.combo.file <- paste0(outdir,"model/",mgs.rf.tag,".combo.train.RData")
save(choose.mgs.prf,train.phe,PDPR.train.phe,file = "../../Result/RandomForest/NPCpairmatch/model/tmp.RData")

#load(paste0("../../Result/RandomForest/NPCpairmatch/model/tmp.RData")) 

if(file.exists(MGS.combo.file)){
  load(MGS.combo.file)
}else{
  mgs.SD2.train <- train.module3(
    choose.mgs.prf,train.phe,
    #"DNAID","SD2",size=8,max.cv=0.2,max.marker=30,maxtry=1,
    "DNAID","SD2",size=4,max.cv=0.3,max.marker=30,maxtry=5,
    thread=60,rept=999,cv.fold=8,step=1.1,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  mgs.DPRP.train <- train.module3(
    choose.mgs.prf,PDPR.train.phe,
    #"DNAID","BE3",size=8,max.cv=0.2,max.marker=30,maxtry=1,
    "DNAID","BE3",size=4,max.cv=0.3,max.marker=30,maxtry=5,
    thread=60,rept=999,cv.fold=8,step=1.1,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  save(mgs.SD2.train, mgs.DPRP.train, file=MGS.combo.file)
}



#performance visual
mgs.train.rf <- mgs.SD2.train$rf
plot.f1(mgs.SD2.train)
mgs.DPPR.train.rf <- mgs.DPRP.train$rf
plot.f1(mgs.DPRP.train)
#step2: train set performance

mgs.tr.mtx <- t(choose.mgs.prf[,which(colnames(choose.mgs.prf)%in%train.phe$DNAID)])
rownames(M0.phe) <- M0.phe$DNAID
mgs.tr.df <- comb.rf.model.pred(mgs.train.rf,mgs.DPPR.train.rf,mgs.tr.mtx,M0.phe)
mgs.tr.df.pred <- mgs.tr.df$pred
mgs.tr.df.auc <- mgs.tr.df$auc


mgs.p1 <- ggtern(data=mgs.tr.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="mgs training set")
mgs.p1

mgs.p3 <- ggplot(mgs.tr.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(mgs.tr.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
mgs.p3

#step3: test set performance

mgs.test.mtx <- t(choose.mgs.prf[,-which(colnames(choose.mgs.prf)%in%train.phe$DNAID)])
mgs.test.df <- comb.rf.model.pred(mgs.train.rf,mgs.DPPR.train.rf,mgs.test.mtx,M0.phe)
mgs.tt.df.pred <- mgs.test.df$pred
mgs.tt.df.auc <- mgs.test.df$auc

mgs.p2 <- ggtern(data=mgs.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="mgs testing set")
mgs.p2

mgs.p4 <- ggplot(mgs.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(mgs.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
mgs.p4

```

**added: Validated by Caucasian cohort
```{r}
#step3: test set performance
ca.phe <- pub.phe%>%filter(cohort=="Caucasian"&CTgrp=="baseline"&Tumor_Type=="NSCLC"&ATB==F&BestEvaluation!="DD")
levels(ca.phe$BestEvaluation) <- c("PD","PD","PR","SD")
ca.phe$BE3 <- droplevels(ca.phe$BestEvaluation)
rownames(ca.phe) <- ca.phe$DNAID
ca.mgs.test.mtx <- t(pub.MGS2.prf[,which(colnames(pub.MGS2.prf)%in%ca.phe$DNAID)])
ca.mgs.test.df <- comb.rf.model.pred(mgs.train.rf,mgs.DPPR.train.rf,ca.mgs.test.mtx,ca.phe)
ca.mgs.tt.df.pred <- ca.mgs.test.df$pred
ca.mgs.tt.df.auc <- ca.mgs.test.df$auc

ca.mgs.p2 <- ggtern(data=ca.mgs.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="ca.mgs testing set")
ca.mgs.p2

ca.mgs.p4 <- ggplot(ca.mgs.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(ca.mgs.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
ca.mgs.p4

```

### KO only

```{r}
#Ms0
ko.rf.tag <- "Ms0.f8_00.ko.BE3"
#select.markers
load("../../Result/ReportScore/MsSign_NPC_BE3/select.markers.RData")
ko.marker.name <- unique(as.character(select.markers$KO))
#M0RAND
#Ms0
##ko.rf.tag <- "MsMd.f8_00.ko.BE3"
##ko.rf.tag <- "MsMd.f8_01.ko.BE3" #same, for check
##load("../../Result/ReportScore/M0RAND_NPC_BE3/select.markers.RData")
#ko.marker.name <- unique(as.character(select.markers$KO))
ko.freq <- ddply(select.markers,"KO",summarise,freq=length(which(table(VS)>0)))
ko.marker.name<- as.character((ko.freq%>%filter(freq>0))$KO)

###
choose.ko.prf <- ko.prf[ko.marker.name,sort(colnames(ko.prf))]

ko.rf.dat <- merge(data.frame(DNAID=colnames(choose.ko.prf),t(choose.ko.prf)),
                train.phe[,c("DNAID","BE2")],by="DNAID")
rownames(ko.rf.dat) <- ko.rf.dat$DNAID
ko.rf.dat <- ko.rf.dat[,-1]

#PDPR
PDPR.ko.rf.dat <- merge(data.frame(DNAID=colnames(choose.ko.prf),t(choose.ko.prf)),
    PDPR.train.phe[,c("DNAID","BE2")],by="DNAID")
rownames(PDPR.ko.rf.dat) <- PDPR.ko.rf.dat$DNAID
PDPR.ko.rf.dat <- PDPR.ko.rf.dat[,-1]

ko.combo.file <- paste0(outdir,"model/",ko.rf.tag,".combo.train.RData")
save(choose.ko.prf,train.phe,PDPR.train.phe,file = "../../Result/RandomForest/NPCpairmatch/model/KOtmp.RData")
if(file.exists(ko.combo.file)){
  load(ko.combo.file) # in cluster
}else{
  #
  ko.SD2.train <- train.module3(
    choose.ko.prf,train.phe,
    #"DNAID","SD2",size=8,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    "DNAID","SD2",size=4,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    thread=60,rept=999,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  ko.DPRP.train <- train.module3(
    choose.ko.prf,PDPR.train.phe,
    #"DNAID","BE3",size=8,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    "DNAID","SD2",size=4,max.cv=0.3,max.marker=30,maxtry=5,ntree=500,
    thread=60,rept=999,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  save(ko.SD2.train, ko.DPRP.train, file=ko.combo.file)
}

#performance visual

ko.train.rf <- ko.SD2.train$rf
plot.f1(ko.SD2.train)
ko.DPPR.train.rf <- ko.DPRP.train$rf
plot.f1(ko.DPRP.train)
#####
#step2: train set performance

ko.tr.mtx <- t(choose.ko.prf[,which(colnames(choose.ko.prf)%in%train.phe$DNAID)])
rownames(M0.phe) <- M0.phe$DNAID
ko.tr.df <- comb.rf.model.pred(ko.train.rf,ko.DPPR.train.rf,ko.tr.mtx,M0.phe)
ko.tr.df.pred <- ko.tr.df$pred
ko.tr.df.auc <- ko.tr.df$auc


ko.p1 <- ggtern(data=ko.tr.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="ko training set")
ko.p1

ko.p3 <- ggplot(ko.tr.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(ko.tr.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
ko.p3

#step3: test set performance

ko.test.mtx <- t(choose.ko.prf[,-which(colnames(choose.ko.prf)%in%train.phe$DNAID)])
ko.test.df <- comb.rf.model.pred(ko.train.rf,ko.DPPR.train.rf,ko.test.mtx,M0.phe)
ko.tt.df.pred <- ko.test.df$pred
ko.tt.df.auc <- ko.test.df$auc

ko.p2 <- ggtern(data=ko.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="ko testing set")
ko.p2

ko.p4 <- ggplot(ko.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(ko.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
ko.p4

```
**added: Validated by Caucasian cohort
```{r}
#step3: test set performance
ca.phe <- pub.phe%>%filter(cohort=="Caucasian"&CTgrp=="baseline"&Tumor_Type=="NSCLC"&ATB==F&BestEvaluation!="DD")
levels(ca.phe$BestEvaluation) <- c("PD","PD","PR","SD")
ca.phe$BE3 <- droplevels(ca.phe$BestEvaluation)
rownames(ca.phe) <- ca.phe$DNAID
ca.ko.test.mtx <- t(pub.KO.prf[,which(colnames(pub.KO.prf)%in%ca.phe$DNAID)])
ca.ko.test.df <- comb.rf.model.pred(ko.train.rf,ko.DPPR.train.rf,ca.ko.test.mtx,ca.phe)
ca.ko.tt.df.pred <- ca.ko.test.df$pred
ca.ko.tt.df.auc <- ca.ko.test.df$auc

ca.ko.p2 <- ggtern(data=ca.ko.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="ca.ko testing set")
ca.ko.p2

ca.ko.p4 <- ggplot(ca.ko.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(ca.ko.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
ca.ko.p4

```


# combine mgs and ko

```{r}
cmk.prf <- rbind(choose.mgs.prf,choose.ko.prf)
cmk.rf.dat <- merge(data.frame(DNAID=colnames(choose.ko.prf),t(choose.mgs.prf),t(choose.ko.prf)),
                train.phe[,c("DNAID","SD2")],by="DNAID")
rownames(cmk.rf.dat) <- cmk.rf.dat$DNAID
cmk.rf.dat <- cmk.rf.dat[,-1]

cmk.rf.tag <- "MsMd.f8_00.cmk.BE3" # contained wilcox MGS markers
cmk.rf.tag <- "MsMd.f8_01.cmk.BE3" # contained anova MGS markers
###
#DPPR
PDPR.cmk.rf.dat <- merge(data.frame(DNAID=colnames(cmk.prf),t(cmk.prf)),
    PDPR.train.phe[,c("DNAID","BE2")],by="DNAID")
rownames(PDPR.cmk.rf.dat) <- PDPR.cmk.rf.dat$DNAID
PDPR.cmk.rf.dat <- PDPR.cmk.rf.dat[,-1]

save(cmk.prf,train.phe,PDPR.train.phe,file = "../../Result/RandomForest/NPCpairmatch/model/combinetmp.RData")
#####CV
cmk.combo.file <- paste0(outdir,"/model/",cmk.rf.tag,".combo.train.RData")
if(file.exists(cmk.combo.file)){
  load(cmk.combo.file)
}else{
  #
  cmk.SD2.train <- train.module3(
    cmk.prf,train.phe,
    #"DNAID","SD2",size=8,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    "DNAID","SD2",size=4,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    thread=60,rept=999,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  cmk.DPRP.train <- train.module3(
    cmk.prf,PDPR.train.phe,
    #"DNAID","BE3",size=8,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    #"DNAID","BE3",size=4,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    "DNAID","BE3",size=4,max.cv=0.2,max.marker=30,maxtry=1,ntree=500,
    #thread=60,rept=999,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    thread=60,rept=999,cv.fold=5,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  save(cmk.SD2.train, cmk.DPRP.train, file=cmk.combo.file)
}

#performance visual
cmk.train.rf <- cmk.SD2.train$rf
plot.f1(cmk.SD2.train)
cmk.DPPR.train.rf <- cmk.DPRP.train$rf
plot.f1(cmk.DPRP.train)
#####

#step2: train set performance

cmk.tr.mtx <- t(cmk.prf[,which(colnames(cmk.prf)%in%train.phe$DNAID)])
rownames(M0.phe) <- M0.phe$DNAID
cmk.tr.df <- comb.rf.model.pred(cmk.train.rf,cmk.DPPR.train.rf,cmk.tr.mtx,M0.phe)
cmk.tr.df.pred <- cmk.tr.df$pred
cmk.tr.df.auc <- cmk.tr.df$auc


cmk.p1 <- ggtern(data=cmk.tr.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cmk training set")
cmk.p1

cmk.p3 <- ggplot(cmk.tr.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(cmk.tr.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
cmk.p3

#step3: test set performance

cmk.test.mtx <- t(cmk.prf[,-which(colnames(cmk.prf)%in%train.phe$DNAID)])
cmk.test.df <- comb.rf.model.pred(cmk.train.rf,cmk.DPPR.train.rf,cmk.test.mtx,M0.phe)
cmk.tt.df.pred <- cmk.test.df$pred
cmk.tt.df.auc <- cmk.test.df$auc

cmk.p2 <- ggtern(data=cmk.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cmk testing set")
cmk.p2

cmk.p4 <- ggplot(cmk.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(cmk.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
cmk.p4

```

```{r}
#ggsave(cmk.p0,width=8,height=4,file=paste0("../../Result/RandomForest/model/",cmk.rf.tag,".NPC.train.auc.raw.plot.pdf"))
ggsave(cmk.p1,width=8,height=8,file=paste0("../../Result/RandomForest/model/",cmk.rf.tag,".NPC.train.ggtern.plot.pdf"))
ggsave(cmk.p2,width=8,height=8,file=paste0("../../Result/RandomForest/model/",cmk.rf.tag,".NPC.test.ggtern.plot.pdf"))
ggsave(cmk.p3,width=5,height=4,file=paste0("../../Result/RandomForest/model/",cmk.rf.tag,".NPC.train.auc.plot.pdf"))
ggsave(cmk.p4,width=5,height=4,file=paste0("../../Result/RandomForest/model/",cmk.rf.tag,".NPC.test.auc.plot.pdf"))

```


# clinical preperation
```{r}
HLA.df <- HLA.mdat[,c("PID","TMBnew","HLAe.SNP")]
HLA.df$HLA.E <- as.numeric(as.factor(HLA.mdat$HLAe.SNP)) -1 
HLA.info<- read.csv("../../Result/HLA/Curated.HLA.TMB.20190101.csv")
#HLA.mdat
#load("../../Result/Phenotype/HLA.mdat.RData")
#load("../../Result/Phenotype/NPC/FecalDNA.Rmd.RData")

#HLA.mdat$ATB00<-F
#HLA.mdat$ATB00 <- ifelse(HLA.mdat$ATB=="baseline",T,F)
#HLA.mdat$ATB00[which(is.na(HLA.mdat$ATB00))] <- F

m2.phe <- merge(M0.phe,HLA.df,by="PID")
m0.phe <- merge(M0.phe,HLA.df,by="PID",all.x=T)
m0.phe$TMB <- m0.phe$TMBnew
m0.phe <- ddply(m0.phe,"Gender",transform,
                TMB.fill=ifelse(is.na(TMB),median(TMB[which(!is.na(TMB))]),TMB),
                HLA.E.fill=ifelse(is.na(HLA.E),median(HLA.E[which(!is.na(HLA.E))]),HLA.E))
A0.phe <- merge(FDNA.NPC.pick3%>%filter(Ms=="M0"),HLA.df,by="PID",all.x=T)
A0.phe$TMB <- A0.phe$TMBnew
A0.phe <- ddply(A0.phe,"Gender",transform,
                TMB.fill=ifelse(is.na(TMB),median(TMB[which(!is.na(TMB))]),TMB),
                HLA.E.fill=ifelse(is.na(HLA.E),median(HLA.E[which(!is.na(HLA.E))]),HLA.E))

A0.PID <- as.character(A0.phe$PID)

PR.train.PID <- as.character(sort((A0.phe%>%filter(BE3=="PR"))$PID)[1:8])

merge.SD.dat <- rbind(
  cbind(set=1,A0.phe%>%filter(PID%in%PR.train.PID)),
  cbind(set=0,A0.phe%>%filter(BE3=="SD")))
#mat.SD <- pairmatch(set~Treatment+Gender+Age+BMI,controls = 1, data=merge.SD.dat)
mat.SD <- pairmatch(set~Treatment+Gender+Age+E,controls = 1, data=merge.SD.dat)

merge.PD.dat <- rbind(
  cbind(set=1,A0.phe%>%filter(PID%in%PR.train.PID)),
  cbind(set=0,A0.phe%>%filter(BE3=="PD")))
#mat.PD <- pairmatch(set~Treatment+Gender+Age+BMI,controls = 1, data=merge.PD.dat)
mat.PD <- pairmatch(set~Treatment+Gender+Age,controls = 1, data=merge.PD.dat)

A0.merge.dat <- unique(rbind(
  merge.SD.dat[which(!is.na(mat.SD)),],
  merge.PD.dat[which(!is.na(mat.PD)),]))

A0.train.PID <- as.character(A0.merge.dat$PID)

A0.train.phe <- A0.phe%>%filter(PID%in%A0.train.PID)
A0.test.phe  <- A0.phe%>%filter(!PID%in%A0.train.PID)

#
m0.train.phe <- m0.phe%>%filter(PID%in%train.PID)
m0.test.phe  <- m0.phe%>%filter(!PID%in%train.PID)

#
table(A0.merge.dat$BE3,A0.merge.dat$Treatment)
table(A0.merge.dat$BE3,A0.merge.dat$Gender)
table(A0.merge.dat$BE3,A0.merge.dat$ATB02)
ggplot(merge.dat,aes(x=Age,color=BE3,linetype=Gender)) + 
  geom_density() + theme_bw()
#ggplot(merge.dat,aes(x=BMI,color=BE3,linetype=Gender)) + 
ggplot(merge.dat,aes(x=E,color=BE3,linetype=Gender)) + 
  geom_density() + theme_bw()
#ggplot(merge.dat,aes(x=Age,y=BMI,color=BE3,shape=Gender)) + 
ggplot(merge.dat,aes(x=Age,y=E,color=BE3,shape=Gender)) + 
 #stat_density2d(h = c(5,10),n=10,alpha=.3) +
  stat_ellipse(aes(group=BE3)) +  
  geom_point() + theme_bw()

```
**clnical model (m0)**

```{r}
#HLA.mdat$EGFR[which(is.na(HLA.mdat$EGFR))] <- "-"
#HLA.mdat$KRAS[which(is.na(HLA.mdat$KRAS))] <- "-"
#HLA.mdat$ALK[which(is.na(HLA.mdat$ALK))] <- "-"
rownames(HLA.mdat) <- HLA.mdat$PID
HLA.mdat$BE3<-HLA.mdat$BestEvaluation
HLA.mdat$SD <- as.factor(ifelse(HLA.mdat$BE3=="SD","SD","PDPR"))
train.HLA.mdat <- HLA.mdat[which(rownames(HLA.mdat)%in%train.phe$PID),]
#test.HLA.mdat <- HLA.mdat[-which(rownames(HLA.mdat)%in%train.phe$PID|HLA.mdat$ATB02==T),] 
test.HLA.mdat <- HLA.mdat[-which(rownames(HLA.mdat)%in%train.phe$PID),]
test2.HLA.mdat <- HLA.info%>%filter(grepl("Q",PID)&Tumor=="NPC"&!is.na(TMB))

train.HLA.mdat <- train.HLA.mdat%>%filter(!is.na(TMBnew))
rownames(train.HLA.mdat) <- train.HLA.mdat$PID
tmp.df <- train.HLA.mdat%>%filter(BE3!="SD")
#tmp.df <- train.HLA.mdat%>%filter(BE3!="SD"&!is.na(TMBnew))
tmp.df$BE3 <- droplevels(tmp.df$BE3)
#train.mod1 <- randomForest(BE3~EGFR+KRAS+ALK+TMB,tmp.df)
#train.mod2 <- randomForest(SD~EGFR+KRAS+ALK+TMB,train.HLA.mdat)
train.mod1 <- randomForest(BE3~TMBnew+TNB,tmp.df)
train.mod2 <- randomForest(SD~TMBnew+TNB,train.HLA.mdat)

#step2: train set performance
HLA.tr.mtx <- train.HLA.mdat
rownames(M0.phe) <- M0.phe$PID
HLA.tr.df <- comb.rf.model.pred(train.mod2,train.mod1,HLA.tr.mtx,M0.phe)
HLA.tr.df.pred <- HLA.tr.df$pred
HLA.tr.df.auc <- HLA.tr.df$auc

HLA.p1 <- ggtern(data=HLA.tr.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="HLA training set")
HLA.p1

HLA.p3 <- ggplot(HLA.tr.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(HLA.tr.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
HLA.p3

#step3: test set performance
HLA.test.df <- comb.rf.model.pred(train.mod2,train.mod1,test.HLA.mdat,M0.phe)
HLA.tt.df.pred <- HLA.test.df$pred
HLA.tt.df.auc <- HLA.test.df$auc

HLA.p2 <- ggtern(data=HLA.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="HLA testing set")
HLA.p2

HLA.p4 <- ggplot(HLA.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(HLA.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
HLA.p4



# combine cmk to HLA
tt.cmk.test.df <- merge(cmk.tt.df.pred,HLA.tt.df.pred[,c("PID","PD","SD","PR")],by="PID",all.x=T)

tt.cmk.test.df$PD <- (tt.cmk.test.df$PD.x * ifelse(is.na(tt.cmk.test.df$PD.y),1,tt.cmk.test.df$PD.y)) * cmk.test.df$conf.PR * HLA.test.df$conf.PR
tt.cmk.test.df$SD <- (tt.cmk.test.df$SD.x * ifelse(is.na(tt.cmk.test.df$SD.y),1,tt.cmk.test.df$SD.y)) * cmk.test.df$conf.SD * HLA.test.df$conf.SD
tt.cmk.test.df$PR <- (tt.cmk.test.df$PR.x * ifelse(is.na(tt.cmk.test.df$PR.y),1,tt.cmk.test.df$PR.y)) * cmk.test.df$conf.PD * HLA.test.df$conf.PD

tmp.sum <- tt.cmk.test.df$PD + tt.cmk.test.df$PR + tt.cmk.test.df$SD
tt.cmk.test.df$PD <- tt.cmk.test.df$PD/tmp.sum
tt.cmk.test.df$SD <- tt.cmk.test.df$SD/tmp.sum
tt.cmk.test.df$PR <- tt.cmk.test.df$PR/tmp.sum

#########################################
tt.cmk.p2 <- ggtern(data=tt.cmk.test.df,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cmk testing set")
tt.cmk.p2
auc.PD <- auc(tt.cmk.test.df$BE3=="PD",tt.cmk.test.df$PD,plot=T)
auc.SD <- auc(tt.cmk.test.df$BE3=="SD",tt.cmk.test.df$SD,plot=T)
auc.PR <- auc(tt.cmk.test.df$BE3=="PR",tt.cmk.test.df$PR,plot=T)

tt.cmk.test.auc.df <- rbind(
  data.frame(pred="PD",auc=as.numeric(auc.PD),attr(auc.PD,"roc")[2],attr(auc.PD,"roc")[3]),
  data.frame(pred="SD",auc=as.numeric(auc.SD),attr(auc.SD,"roc")[2],attr(auc.SD,"roc")[3]),
  data.frame(pred="PR",auc=as.numeric(auc.PR),attr(auc.PR,"roc")[2],attr(auc.PR,"roc")[3])
)
tt.cmk.test.auc.df <- tt.cmk.test.auc.df[rev(order(tt.cmk.test.auc.df$specificities)),]


tt.cmk.p4 <- ggplot(tt.cmk.test.auc.df,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
tt.cmk.p4

if(do.write){
  ggsave(tt.cmk.p2,width=8,height=8,file=paste0("../../Result/RandomForest/model/NPC.cmk_TMB.test.ggtern.plot.pdf"))
  ggsave(tt.cmk.p4,width=5,height=4,file=paste0("../../Result/RandomForest/model/NPC.cmk_TMB.test.auc.plot.pdf"))
}
```

```{r}
#cli.info <- m0.phe[,c("DNAID","BE3","Age","Gender","BMI","Treatment","TMB.fill","HLA.E.fill")]
cli.info <- m0.phe[,c("DNAID","Age","Gender","Treatment","TMB.fill","HLA.E.fill","BestDay","shan","Day","E","TMB.fill","HLA.E.fill")]
rownames(cli.info) <- cli.info$DNAID
cli.info$Treatment <- as.numeric(as.factor(cli.info$Treatment))

cli.rf.tag <- "Ms0.f8_00.cli.BE3"
#cli.rf.tag <- "Ms0.f8_01.cli.BE3" # anova

cli.info$Gender <- as.numeric(cli.info$Gender) -1 
cli.rf.dat <- cli.info[which(rownames(cli.info)%in%as.character(m0.train.phe$DNAID)),-1]
cli.prf <- t(cli.info[,-c(1:2)])
the.train.PID <- as.character(m0.train.phe$PID)
###
#DPPR
PDPR.cli.rf.dat <- merge(data.frame(DNAID=colnames(cli.info),t(cli.info)),
    PDPR.train.phe[,c("DNAID","BE2")],by="DNAID")
rownames(PDPR.cli.rf.dat) <- PDPR.cli.rf.dat$DNAID
PDPR.cli.rf.dat <- PDPR.cli.rf.dat[,-1]

#####CV


cli.combo.file <- paste0(outdir,"/model/",cli.rf.tag,".combo.train.RData")
save(cli.prf,train.phe,PDPR.train.phe,file = "../../Result/RandomForest/NPCpairmatch/model/clitmp.RData")

#load(paste0("../../Result/RandomForest/NPCpairmatch/model/clitmp.RData")) 
#cli.train.phe<-train.phe[,c("DNAID","PID","Age","Gender","Treatment","Thr1st","BestEvaluation","PFS3m","PFS6m","BestDay","PD.Day","BE3","BE2","shan","CTgrp","E","SD2")]

#cli.train.phe<-train.phe[,c("DNAID","PID","Age","Gender","Treatment","Thr1st","BestEvaluation","PFS3m","PFS6m","BestDay","PD.Day","BE3","BE2","shan","CTgrp","E","SD2")]
cli.train.phe<-train.phe[,c("DNAID","duringDetail","Location","Age","Gender","Treatment","Thr1st","BestDay","shan","Day","E")]

#cli.PDPR.train.phe<-PDPR.train.phe[,c("DNAID","PID","Age","Gender","Treatment","Thr1st","BestEvaluation","PFS3m","PFS6m","BestDay","PD.Day","BE3","BE2","shan","CTgrp","E","SD2")]
cli.PDPR.train.phe<-PDPR.train.phe[,c("DNAID","duringDetail","Location","Age","Gender","Treatment","Thr1st","BestDay","shan","Day","E")]


if(file.exists(cli.combo.file)){
  load(cli.combo.file)
}else{
  #
  cli.SD2.train <- train.module3(
    #cli.prf,train.phe,
    cli.prf,cli.train.phe,
    "DNAID","SD2",size=8,max.cv=0.5,max.marker=30,maxtry=5,ntree=500,
    thread=6,rept=80,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  cli.DPRP.train <- train.module3(
    #cli.prf,PDPR.train.phe,
    cli.prf,cli.PDPR.train.phe,
    "DNAID","BE3",size=8,max.cv=0.5,max.marker=30,maxtry=5,ntree=500,
    thread=6,rept=80,cv.fold=8,step=1.2,impGrp=NULL,sam=F,type="FORK",
    candy=NULL,pick=NULL,ex.PID=NULL)
  #
  save(cli.SD2.train, cli.DPRP.train, file=cli.combo.file)
}

#performance visual
cli.train.rf <- cli.SD2.train$rf
plot.f1(cli.SD2.train)
cli.DPPR.train.rf <- cli.DPRP.train$rf
plot.f1(cli.DPRP.train)
#####
cli.train.p <- predict(cli.train.rf, type = "prob")
cli.train.dat <- merge(train.phe,cli.train.p,by.x="DNAID",by.y="row.names")
cli.train.dat <- merge(M0.phe,cli.train.p,by.x="DNAID",by.y="row.names")
auc1 <- auc(cli.train.dat$SD2,cli.train.dat$SD,plot=T)

#step2: pred PD / PR

PDPR.cli.train.p <- predict(cli.DPPR.train.rf, type = "prob")
PDPR.cli.train.dat <- merge(PDPR.train.phe,PDPR.cli.train.p,by.x="DNAID",by.y="row.names")
PDPR.cli.train.dat <- merge(M0.phe,PDPR.cli.train.p,by.x="DNAID",by.y="row.names")
auc2 <- auc(PDPR.cli.train.dat$BE2,PDPR.cli.train.dat$PR,plot=T)


train.auc.df <- rbind(
  data.frame(pred="SD",auc=as.numeric(auc1),attr(auc1,"roc")[2],attr(auc1,"roc")[3]),
  data.frame(pred="PR",auc=as.numeric(auc2),attr(auc2,"roc")[2],attr(auc2,"roc")[3])
)
train.auc.df <- train.auc.df[rev(order(train.auc.df$specificities)),]

ggplot(train.auc.df,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + facet_grid(.~pred) +
  geom_text(data=unique(train.auc.df[,1:2]),aes(x=.9,y=.1,label=paste0("auc=",round(auc,2))))

####
cli.train.p <- comb.rf.model.pred(cli.train.rf,cli.DPPR.train.rf)
cli.train.dat <- merge(train.phe,cli.train.p,by.x="DNAID",by.y="row.names")
cli.train.dat <- merge(M0.phe,cli.train.p,by.x="DNAID",by.y="row.names")

cli.test.mtx <- t(cli.prf[,-which(colnames(cli.prf)%in%cli.train.dat$DNAID)])
cli.test.pred <- comb.rf.model.pred(cli.train.rf,cli.DPPR.train.rf,cli.test.mtx)
cli.test.df <- merge(M0.phe,cli.test.pred,by.x="DNAID",by.y="row.names")

ggtern(data=cli.train.dat,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cli training set")

ggtern(data=cli.test.df,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cli testing set")

```
**clnical model (A0)**
```{r}
#cli.prf <- A0.phe[,c("DNAID","BE3","Age","Gender","BMI","TMB.fill","HLA.E.fill")]
cli.prf <- A0.phe[,c("DNAID","BE3","Age","Gender","TMB.fill","HLA.E.fill","Treatment","BestEvaluation","PFS3m","PFS6m","BestDay","")]
rownames(cli.prf) <- cli.prf$DNAID
#clinical.dat$BMI <- clinical.dat$weight/(clinical.dat$height/100)^2
cli.prf$Gender <- as.numeric(cli.prf$Gender) -1 

cli.rf.dat <- cli.prf[which(rownames(cli.prf)%in%as.character(A0.train.phe$DNAID)),-1]

cli.train.rf <- randomForest(BE3~.,data=cli.rf.dat, importance = T,proximity=T,ntree=5000,mtry=1,nPerm=10)
cli.train.rf
cli.train.p <- predict(cli.train.rf, type = "prob")
cli.train.dat <- merge(A0.train.phe,cli.train.p,by.x="DNAID",by.y="row.names")

cli.test.mtx <- merge(A0.test.phe[,c("DNAID"),F],cli.prf,by="DNAID")

rownames(cli.test.mtx) <- cli.test.mtx$DNAID
cli.test.mtx <- cli.test.mtx[,-1]
cli.test.p <- predict(cli.train.rf,cli.test.mtx, type = "prob")
cli.test.df <- merge(A0.test.phe,cli.test.p,by.x="DNAID",by.y="row.names")

ggtern(data=cli.train.dat,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5) +
  geom_point(size=2) + 
  labs(title="clinical training set")

ggtern(data=cli.test.df,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5) +
  geom_point(size=2) + 
  labs(title="clinical testing set")

```
#hm
```{r}
#
hm.phe <- read.csv("../../Result/Phenotype/export.lymphocytes.20190611.csv")[,c(1:11,16)]
hm0.phe <- hm.phe%>%filter(Timepoint=="M0"&PID!="P006"&!is.na(ALL)&Tumor_Type=="NPC")
rownames(hm0.phe) <- as.character(hm0.phe$PID)
hm0.phe <- merge(hm0.phe,m0.phe[,c("PID","BE3")])
#hm0.phe <- hm0.phe[,-c(1:3)]
train.hm0.mdat <- HLA.mdat[which(HLA.mdat$PID%in%train.phe$PID),]
test.hm0.mdat <- HLA.mdat[-which(HLA.mdat$PID%in%train.phe$PID|HLA.mdat$ATB02==T),]

tmp.df <- train.hm0.mdat[which(train.hm0.mdat$BE3!="SD"),]
tmp.df$BE3 <- droplevels(tmp.df$BE3)
train.mod1 <- randomForest(BE3~EGFR+KRAS+TMB,tmp.df)
train.mod2 <- randomForest(SD~EGFR+KRAS+TMB,train.hm0.mdat)

#
#step2: train set performance
rownames(M0.phe) <- M0.phe$PID
hm0.tr.df <- comb.rf.model.pred(train.mod2,train.mod1,train.hm0.mdat,M0.phe)
hm0.tr.df.pred <- hm0.tr.df$pred
hm0.tr.df.auc <- hm0.tr.df$auc

hm0.p1 <- ggtern(data=hm0.tr.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="hm0 training set")
hm0.p1

hm0.p3 <- ggplot(hm0.tr.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(hm0.tr.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
hm0.p3

#step3: test set performance
hm0.test.df <- comb.rf.model.pred(train.mod2,train.mod1,test.hm0.mdat,M0.phe)
hm0.tt.df.pred <- hm0.test.df$pred
hm0.tt.df.auc <- hm0.test.df$auc

hm0.p2 <- ggtern(data=hm0.tt.df.pred,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="hm0 testing set")
hm0.p2

hm0.p4 <- ggplot(hm0.tt.df.auc,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(data=unique(hm0.tt.df.auc[,1:2]),aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
hm0.p4
#

tt.hm0.test.df <- merge(cmk.tt.df.pred,hm0.tt.df.pred[,c("PID","PD","SD","PR")],by="PID",all.x=T)
tt.hm0.test.df$PD <- (tt.hm0.test.df$PD.x * ifelse(is.na(tt.hm0.test.df$PD.y),1,tt.hm0.test.df$PD.y * hm0.test.df$conf.PR)) * cmk.test.df$conf.PR
tt.hm0.test.df$SD <- (tt.hm0.test.df$SD.x * ifelse(is.na(tt.hm0.test.df$SD.y),1,tt.hm0.test.df$SD.y * hm0.test.df$conf.SD)) * cmk.test.df$conf.SD
tt.hm0.test.df$PR <- (tt.hm0.test.df$PR.x * ifelse(is.na(tt.hm0.test.df$PR.y),1,tt.hm0.test.df$PR.y * hm0.test.df$conf.PD)) * cmk.test.df$conf.PD

###
tmp.sum <- tt.hm0.test.df$PD + tt.hm0.test.df$PR + tt.hm0.test.df$SD
tt.hm0.test.df$PD <- tt.hm0.test.df$PD/tmp.sum
tt.hm0.test.df$SD <- tt.hm0.test.df$SD/tmp.sum
tt.hm0.test.df$PR <- tt.hm0.test.df$PR/tmp.sum
###
tt.hm0.p2 <- ggtern(data=tt.hm0.test.df,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cmk testing set")
tt.hm0.p2
auc.PD <- auc(tt.hm0.test.df$BE3=="PD",tt.hm0.test.df$PD,plot=T)
auc.SD <- auc(tt.hm0.test.df$BE3=="SD",tt.hm0.test.df$SD,plot=T)
auc.PR <- auc(tt.hm0.test.df$BE3=="PR",tt.hm0.test.df$PR,plot=T)

tt.hm0.test.auc.df <- rbind(
  data.frame(pred="PD",auc=as.numeric(auc.PD),attr(auc.PD,"roc")[2],attr(auc.PD,"roc")[3]),
  data.frame(pred="SD",auc=as.numeric(auc.SD),attr(auc.SD,"roc")[2],attr(auc.SD,"roc")[3]),
  data.frame(pred="PR",auc=as.numeric(auc.PR),attr(auc.PR,"roc")[2],attr(auc.PR,"roc")[3])
)
tt.hm0.test.auc.df <- tt.hm0.test.auc.df[rev(order(tt.hm0.test.auc.df$specificities)),]

tt.hm0.p4 <- ggplot(tt.hm0.test.auc.df,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
tt.hm0.p4

```

#Adding
```{r}
# MGS + KO + HLA + HM
t.hm0.tt.df.pred <- hm0.tt.df.pred
colnames(t.hm0.tt.df.pred)[c(40,42,44)] <- c("SD.z","PD.z","PR.z")
tt.mkhh.test.df <- merge(tt.cmk.test.df,t.hm0.tt.df.pred[,c("PID","SD.z","PD.z","PR.z")],by="PID",all.x=T)
tt.mkhh.test.df$PD <- cmk.test.df$conf.PD * tt.mkhh.test.df$PD.x * 
  ifelse(is.na(tt.mkhh.test.df$PD.y),1,tt.mkhh.test.df$PD.y * HLA.test.df$conf.PD) * 
  ifelse(is.na(tt.mkhh.test.df$PD.z),1,tt.mkhh.test.df$PD.z * hm0.test.df$conf.PD)
tt.mkhh.test.df$SD <- cmk.test.df$conf.SD * tt.mkhh.test.df$SD.x * 
  ifelse(is.na(tt.mkhh.test.df$SD.y),1,tt.mkhh.test.df$SD.y * HLA.test.df$conf.SD) * 
  ifelse(is.na(tt.mkhh.test.df$SD.z),1,tt.mkhh.test.df$SD.z * hm0.test.df$conf.SD)
tt.mkhh.test.df$PR <- cmk.test.df$conf.PR * tt.mkhh.test.df$PR.x * 
  ifelse(is.na(tt.mkhh.test.df$PR.y),1,tt.mkhh.test.df$PR.y * HLA.test.df$conf.PR) * 
  ifelse(is.na(tt.mkhh.test.df$PR.z),1,tt.mkhh.test.df$PR.z * hm0.test.df$conf.PR)
###
tmp.sum <- tt.mkhh.test.df$PD + tt.mkhh.test.df$PR + tt.mkhh.test.df$SD
tt.mkhh.test.df$PD <- tt.mkhh.test.df$PD/tmp.sum
tt.mkhh.test.df$SD <- tt.mkhh.test.df$SD/tmp.sum
tt.mkhh.test.df$PR <- tt.mkhh.test.df$PR/tmp.sum
###
tt.mkhh.p2 <- ggtern(data=tt.mkhh.test.df,aes(x=PR,y=PD,z=SD,color=BE3)) + 
  theme_rgbw() + geom_density_tern(aes(color=BE3),bins=5,alpha=.5,linetype=2) +
  geom_point(size=2,alpha=.7) + 
  labs(title="cmk testing set")
tt.mkhh.p2
auc.PD <- auc(tt.mkhh.test.df$BE3=="PD",tt.mkhh.test.df$PD,plot=T)
auc.SD <- auc(tt.mkhh.test.df$BE3=="SD",tt.mkhh.test.df$SD,plot=T)
auc.PR <- auc(tt.mkhh.test.df$BE3=="PR",tt.mkhh.test.df$PR,plot=T)

tt.mkhh.test.auc.df <- rbind(
  data.frame(pred="PD",auc=as.numeric(auc.PD),attr(auc.PD,"roc")[2],attr(auc.PD,"roc")[3]),
  data.frame(pred="SD",auc=as.numeric(auc.SD),attr(auc.SD,"roc")[2],attr(auc.SD,"roc")[3]),
  data.frame(pred="PR",auc=as.numeric(auc.PR),attr(auc.PR,"roc")[2],attr(auc.PR,"roc")[3])
)
tt.mkhh.test.auc.df <- tt.mkhh.test.auc.df[rev(order(tt.mkhh.test.auc.df$specificities)),]

tt.mkhh.p4 <- ggplot(tt.mkhh.test.auc.df,aes(x=1-specificities,y=sensitivities,color=pred)) + 
  geom_abline(linetype=2,color="grey50") + geom_step() + 
  geom_text(aes(x=.9,y=as.numeric(pred)/30,label=paste0("auc=",round(auc,2)))) + theme_bw()
tt.mkhh.p4

```

#stat
```{r}
stat.auc<-rbind(
  data.frame(set="MGS",unique(mgs.tt.df.auc[,1:2])),
  data.frame(set="KO",unique(ko.tt.df.auc[,1:2])),
  data.frame(set="MGS+KO",unique(cmk.tt.df.auc[,1:2])),
  data.frame(set="TMB",unique(HLA.tt.df.auc[,1:2])),
  data.frame(set="MGS+KO+TMB",unique(tt.cmk.test.auc.df[,1:2])),
  data.frame(set="lymphocyte",unique(hm0.tt.df.auc[,1:2])),
  data.frame(set="MGS+KO+lymphocyte",unique(tt.hm0.test.auc.df[,1:2])),
  data.frame(set="MGS+KO+TMB+lymphocyte",unique(tt.mkhh.test.auc.df[,1:2]))
)

p.stat <- ggplot(stat.auc,aes(x=set,y=auc,fill=set)) + geom_bar(stat="identity") +
  facet_grid(pred~.) + coord_flip() + scale_y_continuous(breaks=seq(0,1,0.2))
p.stat
if(do.write){
  ggsave(p.stat,,width=8,height=4,
         filename = "../../Result/RandomForest/model/NPC.multi.combined.AUC.barplot.pdf")
}
```
