---
title: "Compare significance"
output:
  html_notebook:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(plyr)
library(dplyr)
library(reshape2)
library(ggplot2)
library(RColorBrewer)
library(pheatmap)
library("survminer")
library(parallel)
source("TEST.functions.R")

# Global variance for write or not
do.write <- F
# Outpu dir
outdir <- "../../Result/TEST/timeSeries/"
if(do.write&!dir.exists(outdir)){dir.create(outdir)}
# tag
tag <- "NPC.ATBF.times"
```

# 1. Load data

> 1). Phenotype info
>
> 2). 833 samples Speices profile table (MetaPhlAn2)
>
>   1. Only baseline
>
>   2. Only treatment 1 month
>
>   3. Only treatment 2 month
>
>   4. Only treatment 3 month

By default, the script will only print significant records without writing all results.

> Reading V2 profile and phenotype

```{r, warning=FALSE}
# Load thr1st,patSum,pat3M,Nthr,evaDat,pat3M0,pat3M1
load("../../Result/Phenotype/NPC/phenotype.NPC.Rmd.RData")
# Load patDNA0,patDNA3
load("../../Result/Phenotype/NPC/FecalDNA.Rmd.RData")
#tsALL.phe
load("../../Result/TEST/timeSeries/tsALL.phe.RData")
#TMB
TMB.info <- read.csv("../../Result/Phenotype/curated.HLA-E.info.csv")

#FDNA.NS,FDNA.NPC.pick2,FDNA.NPC.pick3
load("../../Result/TimeSeries/NPC/FDNA.NPC.picks.RData")
FDNA.NPC.pick2.ATBF <- FDNA.NPC.pick2%>%filter(ATB02==F)
FDNA.NPC.pick3.ATBF <- FDNA.NPC.pick3%>%filter(ATB02==F)
sp.prf <- read.table("../../Result/01.Profile/MGS_V2.profile")
sp.min.abun <- sort(unique(unlist(sp.prf)))[2]
mgs.V2.anno <- read.table("../../Result/01.Profile/00.IGC.MGS.annotation",sep="\t",header=T)
mgs.V2.anno$mgs.V2 <- sub(":",".",mgs.V2.anno$X)
mgs.V2.anno2 <- mgs.V2.anno[,c(8,2,3)]
mgs.V2.LPS.anno <- read.csv("../../Result/01.Profile/MGS.1507.annoatation.LPS.csv")
mgs.V2.LPS.anno$MGS <- sub(":",".",mgs.V2.LPS.anno$MGS)
rownames(mgs.V2.LPS.anno) <- mgs.V2.LPS.anno$MGS
```

# Curation
- Remove subjects with antibiotics usage during baseline (-2month before treatment) to 2nd month after treatment.

**test randomly pick of baseline**
```{r}
all.base.phe <- FDNA.NS%>%filter(Ms=="M0"&ATB02==F)

randPick <- function(d){
  i <- sample(1:nrow(d),1); return(d[i,c("PID","DNAID")])
}

set.rand.time = 100

#for wilcox, the prefix is "rand.base.wilcox"
#for anova,  the prefix is "rand.base.anova"
if(file.exists(paste0("../../Result/TEST/NPC/rand.base.anova.",set.rand.time,".RData"))){
  load(paste0("../../Result/TEST/NPC/rand.base.anova.",set.rand.time,".RData"))
}else{
  #
  rand.pick.base.phe <- NULL
  for(i in 1:set.rand.time){
    set.seed(i)
    rand.pick.base.phe[[i]] <- ddply(all.base.phe,"PID",randPick)
  }
  #
  cl <- makeCluster(5)
  clusterExport(cl,c("aovFun","all.base.phe","rand.pick.base.phe","sp.prf","mgs.V2.anno2"))
  rand.pick.base.test <- parSapply(cl,1:set.rand.time,FUN=function(x){
    library(dplyr);library(plyr);
    aovFun(sp.prf[,as.character(rand.pick.base.phe[[x]]$DNAID)],
      all.base.phe[which(all.base.phe$DNAID%in%rand.pick.base.phe[[x]]$DNAID),],
      "DNAID","BE3",NULL,ex.0=F)
  },simplify = F)
  stopCluster(cl)
  save(rand.pick.base.phe, rand.pick.base.test,
       file=paste0("../../Result/TEST/NPC/rand.base.anova.",set.rand.time,".RData"))
}
# for wilcox use "p.value", anova use "ori.pval"
p.val.stat <- sapply(rand.pick.base.test, "[[", "ori.pval")
occ.stat <- sapply(rand.pick.base.test, "[[", "Occ.")
enrich.stat <- sapply(rand.pick.base.test, "[[", "Enrich")
p.val.stat.df <- data.frame(
  mgs.V2=rand.pick.base.test[[1]]$mgs.V2,
  mean.pval=rowMeans(p.val.stat),
  mean.occ=rowMeans(occ.stat),
  enrich = unlist(apply(enrich.stat,1,function(x) {
    tb <- rev(sort(table(x)))[1]
    if(is.na(tb)){return(NA)}else{return(names(tb))}})),
  sign.freq=apply(p.val.stat,1,function(x) length(which(x<0.05)))
)


adj.pval.stat <- sapply(rand.pick.base.test, "[[", "adj.pval")
adj.pval.stat.df <- data.frame(
  mgs.V2=rand.pick.base.test[[1]]$mgs.V2,
  mean.pval=rowMeans(adj.pval.stat),
  sign.freq=apply(adj.pval.stat,1,function(x) length(which(x<0.05)))
)

sign100.ori.df <- p.val.stat.df%>%filter(sign.freq>=75&mean.pval<0.05)
sign100.adj.df <- adj.pval.stat.df%>%filter(sign.freq>=75&mean.pval<0.05)

mgs.diff <- c(
  setdiff(sign100.ori.df$mgs.V2,sign100.adj.df$mgs.V2),
  setdiff(sign100.adj.df$mgs.V2,sign100.ori.df$mgs.V2)
)
(merge(p.val.stat.df,adj.pval.stat.df,by="mgs.V2",all=T))%>%filter(mgs.V2%in%mgs.diff)
#mgs.common <- intersect(sign100.ori.df$mgs.V2,sign100.adj.df$mgs.V2)
mgs.common <- sign100.ori.df$mgs.V2
export.sign.df <- merge(p.val.stat.df%>%filter(mgs.V2%in%mgs.common),mgs.V2.LPS.anno,by.x="mgs.V2",by.y="MGS")
#
p.val.stat.df$enrich<-factor(p.val.stat.df$enrich,levels=c("PD","SD","PR"))
prp <- ggplot(p.val.stat.df%>%filter(sign.freq>0),aes(y=mean.pval,x=sign.freq,size=mean.occ)) + 
  geom_hline(yintercept = 0.05,linetype=2,color="grey50") + 
  geom_vline(xintercept = 75,linetype=2,color="grey50") + 
  geom_point(alpha=.3,aes(color=enrich),shape=16) + 
  scale_color_brewer(palette = "Set1",na.value = "grey50") + theme_bw() +
  scale_y_log10(breaks=c(0.001,0.01,0.05,0.1,1)) + 
  annotation_logticks(side='l') +
  labs(y="p value",x="frequency of singinicance",size="Occurrence") +
  scale_x_continuous(breaks=seq(0,100,20))
prp
if(do.write){
  ggsave(prp,width = 4,height=6,
         file = "../../Result/TEST/NPC/rand.100.anova.pval.freq.dot.pdf")
  write.csv(export.sign.df,
            file="../../Result/TEST/NPC/rand.100.anova.sign.mgs.df.csv")
}
```

test each phase

Use anova:
```{r}
if(file.exists("../../Result/TEST/NPC/sum.Ws.BE3.anova.test.RData")){
  load("../../Result/TEST/NPC/sum.Ws.BE3.anova.test.RData")
}else{
  FDNA.NPC.pick2.ATBF$Ws <- droplevels(FDNA.NPC.pick2.ATBF$Ws)
  sum.Ws.BE3.test <- NULL
  for(i in levels(FDNA.NPC.pick2.ATBF$Ws)){
    tsi.phe <- FDNA.NPC.pick2.ATBF%>%filter(Ws==i)
    tsi.phe <- tsi.phe[order(interaction(tsi.phe$Ws,tsi.phe$BE3)),]
    rownames(tsi.phe) <- tsi.phe$DNAID
    tsi.prf <- sp.prf[,as.character(tsi.phe$DNAID)]
    tsi.BE3.in0.w <- aovFun(tsi.prf,tsi.phe,"DNAID","BE3",NULL,ex.0=F)
    sum.Ws.BE3.test[[i]] <- tsi.BE3.in0.w
    ggplot(tsi.BE3.in0.w,aes(x=FDR,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ws.",i,".FDR.hist.pdf"))
    ggplot(tsi.BE3.in0.w,aes(x=ori.pval,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ws.",i,".ori.pval.hist.pdf"))
    ggplot(tsi.BE3.in0.w,aes(x=adj.pval,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ws.",i,".adj.pval.hist.pdf"))
  }
  save(sum.Ws.BE3.test,file="../../Result/TEST/NPC/sum.Ws.BE3.anova.test.RData")
}

if(file.exists("../../Result/TEST/NPC/sum.Ms.BE3.anova.test.RData")){
  load("../../Result/TEST/NPC/sum.Ms.BE3.anova.test.RData")
}else{
  sum.Ms.BE3.test <- NULL
  for(i in levels(FDNA.NPC.pick3.ATBF$Ms)){
    tsi.phe <- FDNA.NPC.pick3.ATBF%>%filter(Ms==i)
    tsi.phe <- tsi.phe[order(interaction(tsi.phe$Ms,tsi.phe$BE3)),]
    rownames(tsi.phe) <- tsi.phe$DNAID
    tsi.prf <- sp.prf[,as.character(tsi.phe$DNAID)]
    tsi.BE3.in0.w <- aovFun(tsi.prf,tsi.phe,"DNAID","BE3",NULL,ex.0=F)
    sum.Ms.BE3.test[[i]] <- tsi.BE3.in0.w
    ggplot(tsi.BE3.in0.w,aes(x=FDR,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ms.",i,".FDR.hist.pdf"))
    ggplot(tsi.BE3.in0.w,aes(x=ori.pval,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ms.",i,".ori.pval.hist.pdf"))
    ggplot(tsi.BE3.in0.w,aes(x=adj.pval,fill=Enrich)) + geom_histogram(position="stack") + xlim(c(0,1))
    ggsave(paste0("../../Result/TEST/NPC/Ms.",i,".adj.pval.hist.pdf"))
  }
  save(sum.Ms.BE3.test,file="../../Result/TEST/NPC/sum.Ms.BE3.anova.test.RData")
}
```


**check significant MGSs**
```{r}
sum.Ms.wilcox.df <- rbind(
  cbind(Ms="M0",sum.Ms.BE3.test$M0),
  cbind(Ms="W1",sum.Ms.BE3.test$W1),
  cbind(Ms="M1",sum.Ms.BE3.test$M1),
  cbind(Ms="M2",sum.Ms.BE3.test$M2),
  cbind(Ms="M3",sum.Ms.BE3.test$M3),
  cbind(Ms="M4",sum.Ms.BE3.test$M4)
)

#if anova
sum.Ms.wilcox.df$p.value<-sum.Ms.wilcox.df$ori.pval
#
sum.Ms.wilcox.matrix <- dcast(sum.Ms.wilcox.df%>%filter(Occ.>0.0),mgs.V2~Ms,value.var = "p.value")
#
select.mgs <- (export.sign.df%>%filter(mgs.V2%in%mgs.common&mean.occ>0.0))$mgs.V2
sum.Ms.wilcox.matrix <- dcast(sum.Ms.wilcox.df%>%filter(mgs.V2%in%select.mgs),mgs.V2~Ms,value.var = "p.value")
#
rownames(sum.Ms.wilcox.matrix) <- sum.Ms.wilcox.matrix[,1]
sum.Ms.wilcox.matrix <- sum.Ms.wilcox.matrix[,-1]

sum.Ms.sign.matrix <- sum.Ms.wilcox.matrix[which(sum.Ms.wilcox.matrix[,1]<0.05),]

sum.Ms.sign.matrix <- apply(sum.Ms.sign.matrix,1:2,function(x) ifelse(is.na(x),1,x))

phm0 <- pheatmap(sum.Ms.sign.matrix,cluster_rows = T,cluster_cols = F,
         color=colorRampPalette(brewer.pal(n = 11, name ="RdGy"))(11),
         breaks=rev(c(1,0.7,0.5,0.2,0.1,0.05,0.01,0.005,0.001,0.0005,0))
)

signifFun <- function(d){
  pcheck <- d$p.value[d$Ms=="M0"] <0.05 & (d$p.value[d$Ms=="M2"] <0.05|d$p.value[d$Ms=="M3"] <0.05)
  echeck <- length(unique(d$Enrich[d$Ms%in%c("M0","M2","M3")])) == 1 && !is.na(unique(d$Enrich[d$Ms%in%c("M0","M2","M3")]))
  if(pcheck & echeck){
    return(d[1,])
  }
}

```


```{r}
ggplot(sum.Ms.wilcox.df%>%filter(Pct>20&Occ.>0.2),
       aes(x=ori.pval,color=Ms)) + geom_density()
```


**median value**
```{r}
deficientFun <- function(d,p){
  ind <- which(d[p+1:3]==min(d[p+1:3]))
  if(length(ind)==1){
    d$deficient <- sub("median.","",colnames(d)[p+ind])
  }else{
    d$deficient <- NA
  }
  return(d)
}
#p=9 for 'anova', p=10 for 'wilcox'
sum.Ms.wilcox.df <- ddply(sum.Ms.wilcox.df,c("Ms","mgs.V2"),deficientFun,9)
PD.Ms.median.matrix <- dcast(sum.Ms.wilcox.df,mgs.V2~Ms,value.var = "median.PD")
SD.Ms.median.matrix <- dcast(sum.Ms.wilcox.df,mgs.V2~Ms,value.var = "median.SD")
PR.Ms.median.matrix <- dcast(sum.Ms.wilcox.df,mgs.V2~Ms,value.var = "median.PR")
#
sum.Ms.median.matrix <- cbind(PD.Ms.median.matrix,SD.Ms.median.matrix[,-1],PR.Ms.median.matrix[,-1])
colnames(sum.Ms.median.matrix)[2:19] <- paste0(rep(c("PD","SD","PR"),each=6),colnames(sum.Ms.median.matrix)[2:19])
rownames(sum.Ms.median.matrix) <- sum.Ms.median.matrix[,1]
sum.Ms.median.matrix <- sum.Ms.median.matrix[,-1]

sum.Ms.median.matrix <- apply(sum.Ms.median.matrix,1:2,function(x) ifelse(is.na(x),0,x))
sum.Ms.median.matrix <- apply(sum.Ms.median.matrix,1:2,function(x) ifelse(x==0,-10,log10(x)))
#
PD.Ms.signif.matrix <- dcast(sum.Ms.wilcox.df%>%filter(Enrich=="PD"),
                             mgs.V2~Ms,value.var = "p.value")
SD.Ms.signif.matrix <- dcast(sum.Ms.wilcox.df%>%filter(Enrich=="SD"),
                             mgs.V2~Ms,value.var = "p.value")
PR.Ms.signif.matrix <- dcast(sum.Ms.wilcox.df%>%filter(Enrich=="PR"),
                             mgs.V2~Ms,value.var = "p.value")
#
PD.Ms.deficient.matrix <- dcast(sum.Ms.wilcox.df%>%filter(deficient=="PD"&p.value<0.05),
                             mgs.V2~Ms,value.var = "deficient")
SD.Ms.deficient.matrix <- dcast(sum.Ms.wilcox.df%>%filter(deficient=="SD"&p.value<0.05),
                             mgs.V2~Ms,value.var = "deficient")
PR.Ms.deficient.matrix <- dcast(sum.Ms.wilcox.df%>%filter(deficient=="PR"&p.value<0.05),
                             mgs.V2~Ms,value.var = "deficient")

#
sum.Ms.signif.matrix <- matrix(NA,ncol=ncol(sum.Ms.median.matrix),
                               nrow=nrow(sum.Ms.median.matrix),
                               dimnames=dimnames(sum.Ms.median.matrix))
#
sum.Ms.signif.matrix[PD.Ms.deficient.matrix$mgs.V2,  1:6] <- ifelse(PD.Ms.deficient.matrix[,-1]=="PD","-")
sum.Ms.signif.matrix[SD.Ms.deficient.matrix$mgs.V2, 7:12] <- ifelse(SD.Ms.deficient.matrix[,-1]=="SD","-")
sum.Ms.signif.matrix[PR.Ms.deficient.matrix$mgs.V2,13:18] <- ifelse(PR.Ms.deficient.matrix[,-1]=="PR","-")
#
sum.Ms.signif.matrix[PD.Ms.signif.matrix$mgs.V2,  1:6] <- ifelse(
  PD.Ms.signif.matrix[,-1]<0.05,"*",sum.Ms.signif.matrix[PD.Ms.signif.matrix$mgs.V2,  1:6])
sum.Ms.signif.matrix[SD.Ms.signif.matrix$mgs.V2, 7:12] <- ifelse(
  SD.Ms.signif.matrix[,-1]<0.05,"*",sum.Ms.signif.matrix[SD.Ms.signif.matrix$mgs.V2, 7:12])
sum.Ms.signif.matrix[PR.Ms.signif.matrix$mgs.V2,13:18] <- ifelse(
  PR.Ms.signif.matrix[,-1]<0.05,"*",sum.Ms.signif.matrix[PR.Ms.signif.matrix$mgs.V2,13:18])

#
sum.Ms.signif.matrix <- apply(sum.Ms.signif.matrix,1:2,function(x) ifelse(is.na(x),"",x))

#
phMd <- pheatmap(sum.Ms.median.matrix[rownames(sum.Ms.sign.matrix),],
         cluster_rows = T,cluster_cols = F,
         color=c("#CCCCCC",colorRampPalette(brewer.pal(n = 11, name ="RdYlGn"))(11)),
         gaps_col = c(6,12),
         display_numbers = sum.Ms.signif.matrix[rownames(sum.Ms.sign.matrix),]
)

phm.hrow.label <- data.frame(hclust=cutree(phMd$tree_row,k=4))
phm.hrow.label$hclust <- as.factor(phm.hrow.label$hclust)

phm.row.order <- phMd$tree_row$labels[phMd$tree_row$order]
#supp the FDR
sum.Ms.BE3.test$M0$FDR <- p.adjust(sum.Ms.BE3.test$M0$ori.pval,method="BH")
sum.sign.anno <- merge(sum.Ms.BE3.test$M0%>%filter(mgs.V2%in%phm.row.order),
                       mgs.V2.LPS.anno,by.x="mgs.V2",by.y="MGS")
rownames(sum.sign.anno) <- sum.sign.anno$mgs.V2
sum.sign.anno <- sum.sign.anno[phm.row.order,]

if(do.write){
  write.csv(sum.sign.anno,file="../../Result/TEST/NPC/mgs.anova.sum.ms.sgin.anno.csv")
}
```

**show enriched**
```{r}
pick.mgs <- sum.Ms.wilcox.df%>%filter(Ms=="M0"&mgs.V2%in%phm.row.order&Pct>=5)
pick.mgs.PD <- pick.mgs%>%filter(Enrich=="PD")
order.mgs.PD <- reorder(pick.mgs.PD$mgs.V2,pick.mgs.PD$median.PD,mean)
pick.mgs.SD <- pick.mgs%>%filter(Enrich=="SD")
order.mgs.SD <- reorder(pick.mgs.SD$mgs.V2,pick.mgs.SD$median.SD,mean)
pick.mgs.PR <- pick.mgs%>%filter(Enrich=="PR")
order.mgs.PR <- reorder(pick.mgs.PR$mgs.V2,pick.mgs.PR$median.PR,mean)

visual.mgs <- c(rev(levels(order.mgs.PR)),rev(levels(order.mgs.SD)),rev(levels(order.mgs.PD)))
visual.gap <- cumsum(c(length(levels(order.mgs.PR)),length(levels(order.mgs.SD)),length(levels(order.mgs.PD))))
phMdv <- pheatmap(sum.Ms.median.matrix[visual.mgs,],
         cluster_rows = F,cluster_cols = F,
         color=c("#CCCCCC",colorRampPalette(brewer.pal(n = 11, name ="RdYlGn"))(11)),
         gaps_col = c(6,12),gaps_row=visual.gap,
         display_numbers = sum.Ms.signif.matrix[visual.mgs,]
)
phMdv
if(do.write){
  ggsave(phMdv,width=8,height=3,
       filename="../../Result/TEST/NPC/mgs.anova.Ms.pheatmap.pdf")
}
```

**stat of HSDturkey**
```{r}
pick.HSD.df <- melt(pick.mgs,id.vars = "mgs.V2",measure.vars = c("SD.PD","PR.PD","PR.SD"),variable.name = "comparison",value.name = "p.val" )
ggplot(pick.HSD.df,aes(x=p.val,fill=comparison)) + geom_histogram(position="dodge") + 
  geom_vline(xintercept = 0.05,linetype=2) + theme_bw()
if(do.write){
  ggsave(width=8,height=3,
  filename = "../../Result/TEST/NPC/mgs.anova.HSDturkey.pval.pdf")
}
```
